Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-03-23 tex/nusselt_paper.tex
first cut at the new abstract

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification ($n_{\rho}$), aspect
ratio, and Prandtl 
number (Pr) constant while varying the superadiabaticity
and Ra.  We find that Ma is a strong function of the superadiabaticity
in both 2D and 3D simulations, and that it is a 
weak function of Ra in 2D.
In 2D simulations, the scaling of the Nusselt Number (Nu) and the Reynolds number (Re)
with Ra weaken
once the mean Ma of the solution is unity, but in 3D
Re and Nu seem to have consistent scaling across Ra.
We further find that when the superadiabaticity and Ma are large,
density inversions form in the thermal boundary layers.
%As Ra increases and $\text{Ma} \rightarrow 1$, a scaling 
%of Nu $\propto$ Ra$^{0.45}$ is observed.  
%As Ra increases to a regime where Ma $\geq 1$,
%this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$. 
%In the regime of Ma $\ll 1$, a consistent
%Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the 
%Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

@@

In the widely-studied \RB (RB) problem of incompressible Boussinesq convection, 
a negative temperature gradient leads to convective instability.
In the evolved solution, upflows and downflows are symmetrical, the
temperature in the interior becomes isothermal and
the conductive flux ($\propto \grad T$) approaches 
zero there. 
For compressible convection in a stratified atmosphere, a
negative entropy gradient leads to convective instability.
Early numerical experiments of moderate-to-high Ma compressible convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed that the evolved state differs from the RB case.
Downflow lanes
become fast and narrow, and upflow lanes turn into broad, slow upwellings.
Furthermore, the \emph{entropy} gradient is negated by convection in the interior, and
a significant temperature gradient and conductive flux persist in the presence of
efficient convection.

@@

Solutions were time-evolved until a long time average of the fluxes
showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_{\text{crit}} = \{11.15, 10.06, 10.97, 10.97\}$ 
for $\epsilon = \{1.0, 0.5, 10^{-4}, 10^{-7}\}$, respectively.

@@

The efficiency of convection is quantified by the Nusselt number (Nu).  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number 
\cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv,z}} + F_{\text{cond,z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond,z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv,z}}\rangle}{\langle F_{\text{cond,z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv,z}}$ and $F_{\text{cond,z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
and $\langle \rangle$ are volume averages.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the conductive flux of the proper corresponding adiabatic atmosphere
For a compressible ideal gas in hydrostatic equilibrium,
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ \cite{spiegel&veronis1960}.  
It is important to measure the evolved value of
$\langle \kappa \rangle = \langle \rho\chi \rangle$, which is nearly
$\kappa_0$ when $\epsilon$ is small but can change appreciably for large
values of $\epsilon$.
In incompressible Boussinesq convection, where $\grad S = 0$ only when 
$\grad T = 0$, this definition reduces to the traditional definition
of the Nusselt number \cite{otero&all2002, johnston&doering2009}.

@@

We have observed interesting phenomena, such as time-dependent oscillating shearing ``windy''
states in our 2D simulations.  Unlike in RB \cite{goluskin&all2014}, our solutions with
windy states have transport which scales in line with our non-windy solutions.
Furthermore, the stratification of 
these polytropic atmospheres evolves in a complex
manner.  Future work should aim to 
understand the importance of stratification on
convective heat transport and other flow properties.

>>>>

\begin{abstract}
Convection in astrophysical systems is stratified and
often occurs at high Rayleigh number (Ra) and low
Mach number (Ma).
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We perform a suite of 2D and 3D simulations in which we vary the initial
superadiabaticity ($\epsilon$) and Ra at a set
initial density stratification ($n_{\rho}$), aspect
ratio, and Prandtl number (Pr).
We find that Ma is a strong function of $\epsilon$
in both 2D and 3D simulations, and that it is a 
weak function of Ra in 2D.
The scaling of evolved fluid properties (Nusselt number, Reynolds number,
stratification) differs in the subsonic and supersonic regimes; however,
aside from the sonic transition, the evolved flow properties
of low and high Ma flows are nearly identical.  Our limited 3D runs suggest that
evolved fluid properties behave similarly in 2D and 3D, but with modified
scaling laws.
\end{abstract}
\maketitle

@@

In the widely-studied \RB (RB) problem of incompressible Boussinesq convection, 
a negative temperature gradient causes convective instability.
In the evolved solution, upflows and downflows are symmetrical, the
temperature in the interior becomes isothermal, and
the conductive flux ($\propto \grad T$) approaches 
zero there. 
For compressible convection in a stratified atmosphere, a
negative entropy gradient causes convective instability.
Early numerical experiments of moderate-to-high Ma compressible convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed a different evolved state from the RB case.
Downflow lanes
become fast and narrow, and upflow lanes turn into broad, slow upwellings.
Furthermore, the \emph{entropy} gradient is negated by convection in the interior, so
a significant temperature gradient and conductive flux persist despite
efficient convection.

@@

Solutions were time-evolved until a long time average of the fluxes
showed little
variance with depth. A linear stability analysis determined
that convective onset
occurs at $\text{Ra}_{\text{crit}} = \{11.15, 10.06, 10.97, 10.97\}$ 
for $\epsilon = \{1.0, 0.5, 10^{-4}, 10^{-7}\}$, respectively.

@@

The efficiency of convection is quantified by the Nusselt number (Nu).  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number 
\cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv,z}} + F_{\text{cond,z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond,z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv,z}}\rangle}{\langle F_{\text{cond,z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv,z}}$ and $F_{\text{cond,z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
and $\langle \rangle$ are volume averages.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the conductive flux of the proper corresponding adiabatic atmosphere
For a compressible ideal gas in hydrostatic equilibrium,
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ \cite{spiegel&veronis1960}.  
It is important to measure the evolved value of
$\langle \kappa \rangle = \langle \rho\chi \rangle$, which is nearly
$\kappa_0$ for small $\epsilon$ but changes appreciably for large
values of $\epsilon$.
In incompressible Boussinesq convection, where $\grad S = 0$ only when 
$\grad T = 0$, this definition reduces to the traditionally defined
Nusselt number \cite{otero&all2002, johnston&doering2009}.

@@

We have observed interesting phenomena, such as time-dependent oscillating shearing windy
states in our 2D simulations.  Unlike in RB \cite{goluskin&all2014}, our solutions with
windy states have transport which scales in line with our non-windy solutions.
Furthermore, the stratification of 
these polytropic atmospheres evolves in a complex
manner.  Future work should aim to 
understand the importance of stratification on
convective heat transport and other flow properties.
